### 2020-06-05 report/sections/server/node/limitations.tex
Report: complete chapter 4

However, the developer may choose to propagate values
between states through persistent storage APIs (such as
storing values in a database and retrieving them for
handlers of continuation states), so our generated APIs
do not strictly enforce scoping to be the only way to
propagate values in the application logic.

>>>>

However, the developer may choose to propagate values
between states through persistent storage APIs (such as
storing values in a database and retrieving them for
handlers of continuation states), so our generated APIs
do not \textit{strictly} enforce scoping to be the only way to
propagate values in the application logic.

### 2020-06-05 report/sections/server/node/runtime.tex
Report: complete chapter 4

We define the session runtime for the \trole{Svr} endpoint
of the \tprotocol{Adder} protocol in Svr.ts,
named after the endpoint.
It exposes a \textbf{public API}
(\cref{lst:noderuntimepublicapi})
with seams for the developer to pass in the WebSocket server 
and application logic
(i.e. their handler implementations).
It is the responsibility of the developer to construct the
WebSocket server and set it up to listen for incoming connections.
Internally, it keeps a \textit{private API}
for executing the EFSM, when all participants have 
joined the session.

@@

\begin{figure}[!h]
\begin{lstlisting}[language=javascript,title=Svr.ts]
export class Svr {
	constructor(wss: WebSocket.Server,
				initialState: Implementation.S51) { ... }
	...
}

// Not exported to developer
class Session {
	private wss: WebSocket.Server;
	private initialState: Implementation.S51;
	private roleToSocket: RoleToSocket;
	...
}
\end{lstlisting}
\captionof{lstlisting}{Public and Private API of Session Runtime for Server Endpoint}
\label{lst:noderuntimepublicapi}
\end{figure}

@@

\begin{itemize}
\item use set and partial object to keep track of pending connections
\item ws event listeners
\item forward reference that we need to manage cancellations here too
\end{itemize}

@@

The \texttt{Session} class executes the EFSM.
We define a transition function, \texttt{next()}, 
parameterised by the current state,
which invokes the handler defined by the develoepr
and performs the required channel actions for non-terminal states:

@@

\item Empty queues are initialised for both mappings
in the constructor (\cref{line:queue1,,line:queue2}).

>>>>

We define the session runtime for the \trole{Svr} endpoint
of the \tprotocol{Adder} protocol in Svr.ts,
named after the endpoint.
It exposes a \textbf{public API}
with seams for the developer to pass in the WebSocket server 
and application logic
(i.e. their handler implementations).
It is developer's responsibility to construct the
WebSocket server and set it up to listen for incoming connections.
Internally, it keeps a \textit{private API}
for executing the EFSM, when all participants have 
joined the session.

\begin{lstlisting}[language=javascript]
export class Svr {
	constructor(wss: WebSocket.Server,
				initialState: Implementation.S51) { ... }
	...
}

// Not exported to developer
class Session {
	private wss: WebSocket.Server;
	private initialState: Implementation.S51;
	private roleToSocket: RoleToSocket;
	...
}
\end{lstlisting}

@@

@@

The constructor of the public API class sets up
the framework for mapping incoming WebSocket connections to
participants. 
The main challenge is to wait for all participants to 
connect to the server endpoint before EFSM execution begins.
We address this by defining an internal protocol for managing
session joining -- since we generate the runtime for both server 
and browser endpoints, we can implement this in a way that is 
transparent to the developer.

\begin{figure}[!h]
\begin{lstlisting}[language=javascript,tabsize=2]
const waiting: Set<Roles.Peers> = new Set([Roles.Peers.Client]); (*@\label{line:initwaiting}@*)

const roleToSocket: Partial<RoleToSocket> = { (*@\label{line:initroletosocket}@*)
	[Roles.Peers.Client]: undefined,
};

const onSubscribe = ({ data, target: socket }) => {
	const { connect: role } = JSON.parse(data) 
		as Message.ConnectRequest; (*@\label{line:connectionrequest}@*)
	if (!waiting.has(role)) { return socket.close(); } (*@\label{line:occupied}@*)

	roleToSocket[role] = socket; (*@\label{line:connectwsbind}@*)
	waiting.delete(role);

	if (waiting.size === 0)
		new Session(wss, roleToSocket as RoleToSocket, initialState); (*@\label{line:newsession}@*)
};

wss.addEventListener('connection', ws => {
	ws.onmessage = onSubscribe; (*@\label{line:wslisteneroverride}@*)
});
\end{lstlisting}
\captionof{lstlisting}{\dots}
\label{lst:nodeconnect}
\end{figure}

We show this in \cref{lst:nodeconnect} and walk through
the main parts:

\begin{enumerate}
\item
The server keeps track of the participants that 
have yet to join the session -- this is initialised to the
complete set of
non-server endpoints at the start (\cref{line:initrolesocket}).

\item
Browser endpoints request to join the session
by sending a \textit{connection request} with the role
identifier as payload (\cref{line:connectionrequest}). 
We generate role enums for browser targets
in the same way, so the server can correctly
interpret the message.
We listen to connection requests by overriding the
\texttt{onmessage} event listener for every new connection
(\cref{line:wslisteneroverride}).

\item
If the role is already occupied, then the server
responds by closing the connection (\cref{line:occupied}).

\item
Otherwise, the role is not occupied, so the server binds
the WebSocket (which the message was received from) to the role
(\cref{line:connectwsbind}).
This is accumulated in an interface type mapping each
role to an \textit{optional}\footnote{
TypeScript provides utility types for common type transformations:
\texttt{Partial<T>} constructs a type with all properties
of \texttt{T} set to optional.
} 
WebSocket property
-- for roles who have
not yet connected, we do not know the WebSocket binding, so the
WebSocket for these roles are \lstonelinejs{undefined}.

\item
When the server is no longer waiting for any participants,
it notifies all other roles through the bounded WebSocket
connections that the session will start, and delegates
EFSM execution to the private API by constructing
an instance of the \texttt{Session} class (\cref{line:newsession}).
The notification process is managed by the \texttt{Session} class.

\begin{lstlisting}[language=javascript,tabsize=2,numbers=none]
Object.values(roleToSocket).forEach(socket => {
	socket.send(JSON.stringify(Message.ConnectConfirm));
});
\end{lstlisting}

\item
We define the interfaces and factories for connection messages 
in the generated EFSM.ts file under
the \texttt{Message} namespace.

\begin{lstlisting}[language=javascript,tabsize=2,numbers=none]
// Inside the Message namespace...
export interface ConnectRequest { connect: Role.Peers };
export const ConnectConfirm = { connected: true };
\end{lstlisting}
\end{enumerate}

@@

The \texttt{Session} class executes the EFSM.
We define a transition function, \texttt{next()}, 
parameterised by the current state,
The constructor of the \texttt{Session} class
explicitly calls \texttt{next()} with the 
initial state implementation provided by the developer
to start EFSM execution.
\texttt{next()} invokes the handler defined by the developer
and performs the required channel actions for non-terminal states.

@@

\item Empty queues are initialised for both mappings
(\cref{line:queue1,,line:queue2}).
